---
title: "World Development Indicators and Gold Price Data Analysis"
author: "Adrian Glapiński"
date: "`r format(Sys.time(), '%d/%m/%Y')`"
output: 
  html_document: 
    code_folding: "hide"
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, 
                      out.width = "100%")
set.seed(18)
```

## Executive summary
In this report, firstly, all data sets were cleaned and then a comprehensive
summary with statistics was presented for each of them. Then, 

## Libraries used
Following libraries and dependencies were used in the report:

```{r load-libraries, class.source="fold-show"}
library(knitr)

library(kableExtra)
library(DT)

library(tidyverse)
library(readxl)
library(janitor)

library(lubridate)

library(skimr)

library(corrr)

library(ggplot2)
library(ggthemes)
library(gganimate)
library(ggcorrplot)
library(plotly)

library(caret)
library(missForest)

library(countrycode)
```

## Loading data
The input data includes information on the economic indicators and 
development of each country as measured by over 200 statistics. In addition, 
it includes information on currency exchange rates, gold prices, 
Bitcoin trading, and the monthly performance of the S&P Composite. 

The data was collected by various institutions, 
notably the World Bank and Nasdaq.

```{r load-data}
# https://data.worldbank.org/
wdi_df <- read_excel('./data/World_Development_Indicators.xlsx', 
                     na = c('', '..')) %>% head(-5)

currency_df <- read_csv('./data/CurrencyExchangeRates.csv', 
                        col_types = cols(Date = col_date("%Y-%m-%d")))

# https://data.nasdaq.com/data/LBMA/GOLD-gold-price-london-fixing
gold_prices_df <- read_csv('./data/Gold prices.csv')

# https://data.nasdaq.com/data/YALE/SPCOMP-sp-composite
sp_df <- read_csv('./data/S&P Composite.csv')

# https://data.nasdaq.com/data/BCHAIN-blockchain
btc_diff_df <- read_csv('./data/bitcoin/BCHAIN-DIFF.csv')
btc_hrate_df <- read_csv('./data/bitcoin/BCHAIN-HRATE.csv')
btc_mkpru_df <- read_csv('./data/bitcoin/BCHAIN-MKPRU.csv')
btc_trvou_df <- read_csv('./data/bitcoin/BCHAIN-TRVOU.csv')
```

## Cleaning data sets
While cleaning each data set, decisions described below were made.

### World Development Indicators
0. As part of loading data, empty and `..` values were interpreted as `NA`
values. Last 5 rows were also dropped, because they contained data set metadata,
such as last update date.
1. `Country Code` and `Series Code` columns were dropped, as they 
were not useful in further analysis
2. All columns except `Country Name` and `Series Name` were pivoted into 
longer format; names of columns were stored into `Year` column 
3. All columns were pivoted into wider format, taking column names from 
`Series Name` column
4. `Country Name` column was mutated into column of `factor` type and 
`Year` column was mutated into column of `numeric` type
5. Column names were cleaned - all words were transformed into `snake_case`
```{r clean-wdi-data-set}
wdi_cleaned_df <- wdi_df %>%
  select(-c(`Country Code`, `Series Code`)) %>%
  pivot_longer(-c(`Country Name`, `Series Name`), names_to = 'Year') %>%
  pivot_wider(names_from = `Series Name`) %>%
  mutate(`Country Name` = as_factor(`Country Name`),
         Year = as.numeric(word(Year))) 
wdi_names <- names(wdi_cleaned_df) %>% 
  set_names(make_clean_names(.)) %>%
  as.list
wdi_cleaned_df <- set_names(wdi_cleaned_df, names(wdi_names))
```

### Currency Exchange Rates
0. As part of loading data, `Date` column was parsed as column of `date` type
1. All columns except `Date` were pivoted into longer format; names of columns
were stored into `currency` column, values were stored into `exchange_rate` 
column
2. Rows with `NA` values were dropped
3. Column names were cleaned - all words were transformed into `snake_case`
4. `currency` column was mutated into column of `factor` type
```{r clean-currency-data-set}
currency_df <- currency_df %>%
  pivot_longer(cols = -Date, names_to = 'currency',
               values_to = 'exchange_rate') %>%
  drop_na() %>%
  clean_names() %>%
  mutate(currency = as_factor(currency))
```

### Gold prices
1. Column names were cleaned - all words were transformed into `snake_case`
```{r clean-gold-prices-data-set}
gold_prices_df <- clean_names(gold_prices_df)
```

### S&P Composite
1. Column names were cleaned - all words were transformed into `snake_case`
2. `year` column was renamed to `date`
```{r clean-sp-composite-data-set}
sp_df <- clean_names(sp_df) %>% rename(date = year)
```

### Bitcoin
1. Column names were cleaned - all words were transformed into `snake_case`
2. All data frames loaded from CSV files were joined into one data frame
3. Following columns were renamed:
    - `value_diff` to `difficulty`
    - `value_hrate` to `hash_rate`
    - `value_mkpru` to `market_price_usd`
    - `value_trvou` to `exchange_trade_volume_usd`
```{r clean-btc-data-set}
btc_diff_df <- clean_names(btc_diff_df)
btc_hrate_df <- clean_names(btc_hrate_df)
btc_mkpru_df <- clean_names(btc_mkpru_df)
btc_trvou_df <- clean_names(btc_trvou_df)

btc_df <- btc_diff_df %>%
  full_join(btc_hrate_df, by = 'date', suffix = c('_diff', '_hrate')) %>%
  full_join(btc_mkpru_df, by = 'date') %>% 
  full_join(btc_trvou_df, by = 'date', suffix = c('_mkpru', '_trvou')) %>%
  rename(difficulty = value_diff,
         hash_rate = value_hrate,
         market_price_usd = value_mkpru,
         exchange_trade_volume_usd = value_trvou)
```

## Cleaned data set summary
In this section, summary with basic statistics for each of the data sets 
and their variables was presented. 
```{r data-set-summary}
tribble(
  ~Table, ~`No. of columns (attributes)`, ~`No. of rows (observations)`,
  'World Development Indicators', ncol(wdi_cleaned_df), nrow(wdi_cleaned_df),
  'Currency Exchange Rates', ncol(currency_df), nrow(currency_df),
  'Gold prices', ncol(gold_prices_df), nrow(gold_prices_df),
  'S&P Composite', ncol(sp_df), nrow(sp_df),
  'Bitcoin', ncol(btc_df), nrow(btc_df)
) %>% kbl() %>% kable_styling()
```

### World Development Indicators
```{r wdi-summary}
report_table <- function(df, round_digits=2) {
  datatable(df, style = 'bootstrap', rownames = FALSE,
            options = list(scrollX = TRUE)) %>%
    formatRound(names(select_if(df, is.numeric)), round_digits)
}

wdi_skim_df <- wdi_cleaned_df %>% 
  set_names(wdi_names) %>%
  skim() 
wdi_skim_df %>%
  yank('factor') %>% 
  clean_names(case = 'title') %>%
  report_table
wdi_skim_df %>%
  yank('numeric') %>% 
  clean_names(case = 'title') %>%
  report_table()
```

### Currency Exchange Rates
```{r currency-summary}
currency_df %>% 
  clean_names(case = 'title') %>%
  skim() %>% 
  clean_names(case = 'title') %>%
  report_table()
```
A closer look at the data in `Currency Exchange Rates` data set reveals 
ambiguity whether currency exchange rates are from given currency to U.S. Dollar
or vice versa. By looking at exchange rate for Polish Zloty, we can deduce that
exchange rates in the data set are from U.S. Dollar to Polish Zloty.
On the other side, looking at exchange rates of Euro or U.K. Pound Sterling
suggests that the conversion in these cases is reverse; it is common knowledge
that Euro and U.K. Pound Sterling are more valuable currencies than U.S. Dollar, 
so their exchange rate should be below `1.0`, but it isn't.
Therefore, **whole data set was discarded from further analysis**.

```{r}
currency_df %>% 
  filter(currency %in% c('Polish Zloty', 'U.K. Pound Sterling', 'Euro',
                         'U.S. Dollar')) %>%
  group_by(date) %>% 
  filter(all(c('Polish Zloty', 'U.K. Pound Sterling', 'Euro', 'U.S. Dollar')
               %in% currency)) %>% 
  tail(4) %>% 
  clean_names(case = 'title') %>% 
  kbl() %>% kable_styling()
```

### Gold prices
```{r gold-prices-summary}
gold_prices_df %>% 
  clean_names(case = 'title', abbreviations = c('USD', 'GBP', 'AM', 'PM')) %>%
  skim() %>%
  clean_names(case = 'title') %>%
  report_table()
```

### S&P Composite
```{r sp-composite-summary}
sp_df %>% 
  clean_names(case = 'title', abbreviations = c('CPI')) %>% 
  rename(`S&P Composite` = `S p Composite`) %>%
  skim() %>% 
  clean_names(case = 'title') %>%
  report_table()
```

### Bitcoin
```{r bitcoin-summary}
btc_df %>% 
  clean_names(case = 'title', abbreviations = c('USD')) %>%
  skim() %>%
  clean_names(case = 'title') %>%
  report_table()
```

## Finding correlations
Correlations between World Development Indicators were presented on the 
interactive correlation heat map below. **Only variables with more than 50%** 
**complete rate (less than 50% of `NA` values) were selected**.
Correlation was calculated using the `pairwise.complete.obs` option which 
ensures that only complete pairs of observations of variables were used.

Hover over cells in the heat map to show correlated variables names 
and their correlation values.
```{r wdi-correlations}
wdi_na_filtered_df <- wdi_cleaned_df %>%
  set_names(wdi_names) %>%
  select(where(~mean(is.na(.)) < 0.50))

wdi_cor_matrix <- wdi_na_filtered_df %>%
  select(3:last_col()) %>%
  cor(use = 'pairwise.complete.obs')
wdi_cor_matrix[!lower.tri(wdi_cor_matrix)] <- NA

corr_plot <- wdi_cor_matrix %>% 
  ggcorrplot() +
  labs(x = 'Variable 1', y = 'Variable 2') + 
  theme_classic() +
  theme(axis.text = element_blank(), axis.ticks = element_blank()) 

ggplotly(corr_plot)
```

Below the heat map, table with all correlations in range (0.6, 0.9) was also 
included. Pair of variables with correlation higher than 0.9 were excluded,
because they showed obvious relationships between variables, like percentage
of urban and rural population or percentage of male and female population.
```{r}
wdi_cor_matrix %>%
  as_cordf() %>%
  stretch() %>% 
  filter(abs(r) > 0.6, abs(r) < 0.9) %>%
  arrange(desc(abs(r))) %>%
  rename(X = x, Y = y, Correlation = r) %>%
  report_table(round_digits = 5)
```

### GDP per capita correlations
Per capita gross domestic product (GDP) measures a country's economic output per
person and is calculated by dividing the GDP of a country by its population.
Per capita GDP is a global measure used for gauging the prosperity of nations.

```{r}
wdi_map_df <- wdi_cleaned_df %>% 
  filter(!country_name %in% c('Channel Islands', 'Kosovo', 
                              'Low & middle income', 'Low income', 
                              'Lower middle income', 'Middle income', 'World',
                              'Upper middle income', 'High income')) 

wdi_map_df %>%
  rename(`GDP per capita (USD)` = gdp_per_capita_current_us) %>%
  plot_geo(locationmode = 'country names') %>%
  add_trace(z = ~`GDP per capita (USD)`, zmin = 0, zmax = 70000, 
            color = ~`GDP per capita (USD)`,
            frame = ~year, 
            colors = 'Blues',
            text = ~country_name,
            locations = ~country_name,
            marker = list(line = list(color = toRGB("grey"), width = 0.5))) %>%
  colorbar(title = 'GDP per capita (USD)') %>% 
  layout(title = 'GDP per capita over years',
         geo = list(scope = 'world',
                    projection = list(type = 'natural earth'),
                    countrycolor = toRGB('grey'),
                    showcoastlines = TRUE)) %>%  
  animation_opts(redraw = FALSE) %>%  
  plotly_build()
```

#### Correlation between GDP per capita and life expectancy
Preston curve is an empirical relationship between life expectancy 
and GDP per capita. It indicates that individuals born in richer countries, 
on average, can expect to live longer than those born in poor countries. 
The first plot is a reproduction of the Preston curve for year 2019 
and countries with GDP per capita below 70000 USD.
```{r}
wdi_continents_df <- wdi_map_df %>%
  add_column(continent = countrycode(sourcevar = .$country_name,
                                     origin = "country.name",
                                     destination = "continent")) %>%
  relocate(continent)

preston_curve_plot <- wdi_continents_df %>% 
  filter(year == 2019, gdp_per_capita_current_us < 70000) %>%
  ggplot(aes(x = gdp_per_capita_current_us, y = life_expectancy_at_birth_total_years)) +
  geom_point(aes(color = continent, 
                text = paste('Country:', country_name,
                             '<br>Continent:', continent,
                             '<br>GDP per capita (USD):', round(gdp_per_capita_current_us),
                             '<br>Life expectancy (years):', round(life_expectancy_at_birth_total_years, digits = 2)))) +
  labs(x = 'GDP per capita (USD)',
       y = 'Life expectancy (years)', 
       color = "Continent") + 
  geom_smooth(se = FALSE)

ggplotly(preston_curve_plot, tooltip = 'text')
```

The second plot presents how GDP per capita and life expectancy changed over
years in countries grouped by continents. It also shows that in general, 
countries with higher GDP tend to have a higher life expectancy. It shows
that with passing years standard of living is improving across the world
as well.
```{r}
wdi_continents_df %>%
  ggplot(aes(gdp_per_capita_current_us, 
             life_expectancy_at_birth_total_years, 
             size = population_total, 
             color = continent)) +
  geom_point(alpha = 0.7) +
  scale_size_continuous(name = "Population",
                        breaks = c(1e9, 1e8, 1e7, 1e6),
                        limits = c(1e6, 1e9),
                        range = c(1, 8)) +
  scale_x_log10(labels = scales::comma) +
  theme_bw(base_size = 8) +
  labs(title = "Year: {round(frame_time)}", 
       x = 'GDP per capita (USD)',
       y = 'Life expectancy (years)', 
       color = "Continent",
       caption = "") +
  transition_time(year) +
  ease_aes('linear') 
```

High positive correlation can be immediately deduced by observing the 
animation, which is also confirmed by the following map plot.
```{r}
wdi_life_expectancy_cor_df <- wdi_map_df %>% 
  group_by(country_name) %>%
  summarize(correlation = cor(gdp_per_capita_current_us,
                              life_expectancy_at_birth_total_years, 
                              use = 'pairwise.complete.obs'))

wdi_life_expectancy_cor_df %>%
  plot_geo(locationmode = 'country names') %>%
  add_trace(z = ~correlation, zmin = -1, zmax = 1, 
            color = ~correlation,
            colors = c("blue", "white", "red"),
            text = ~country_name,
            locations = ~country_name,
            marker = list(line = list(color = toRGB("grey"), width = 0.5))) %>%
  colorbar(title = 'Correlation') %>% 
  layout(title = 'GDP per capita (USD) - life expectancy (years) correlation',
         geo = list(scope = 'world',
                    projection = list(type = 'natural earth'),
                    countrycolor = toRGB('grey'),
                    showframe = TRUE,
                    showcoastlines = TRUE)) %>%  
  plotly_build()
```

Standard of living is the material well being of the average person in a
given population. Life expectancy increases with the standard of living.
GDP per capita is a good approximation for standard of living,
which explains why there is a high correlation between GDP per capita and 
life expectancy.

#### Correlation between GDP per capita and infant mortality rate
To validate that decrease in per capita GDP is associated with an increase in 
infant mortality rate, as found by the authors of 
[Aggregate Income Shocks and Infant Mortality in the Developing World paper](https://www2.gwu.edu/~iiep/assets/docs/papers/Baird_IIEPWP2010-7.pdf), 
the variables change over years 1999-2015 was presented in selected countries. 
```{r}
infmort_plot <- wdi_continents_df %>%
  filter(country_name %in% c("United States", "Tonga", "Colombia", "Grenada", 
                             "Sri Lanka", "Malta", "Germany", "Japan", "Sweden", 
                             "Netherlands"),
         year %in% 1999:2015) %>%
  rename(`GDP per capita (USD)` = gdp_per_capita_current_us,
         `Infant mortality rate (per 1,000 live births)` =
           mortality_rate_infant_per_1_000_live_births,
         Continent = continent,
         Country = country_name) %>%
  ggplot(aes(x = `GDP per capita (USD)`, 
             y = `Infant mortality rate (per 1,000 live births)`,
             color = Continent,
             tooltip = Country)) + 
  geom_point() +
  facet_wrap(~year)

ggplotly(infmort_plot)
```

Once again, high correlation can be deduced from the plots, but this time 
it is negative, which is confirmed by the following map plot. 
```{r}
wdi_mortality_rate_cor_df <- wdi_map_df %>% 
  group_by(country_name) %>%
  summarize(correlation = cor(gdp_per_capita_current_us,
                              mortality_rate_infant_per_1_000_live_births, 
                              use = 'pairwise.complete.obs'))

wdi_mortality_rate_cor_df %>%
  plot_geo(locationmode = 'country names') %>%
  add_trace(z = ~correlation, zmin = -1, zmax = 1, 
            color = ~correlation,
            colors = c("blue", "white", "red"),
            text = ~country_name,
            locations = ~country_name,
            marker = list(line = list(color = toRGB("grey"), width = 0.5))) %>%
  colorbar(title = 'Correlation') %>% 
  layout(title = 'GDP per capita (USD) - mortality rate, infant (per 1,000 live births) correlation',
         geo = list(scope = 'world',
                    projection = list(type = 'natural earth'),
                    countrycolor = toRGB('grey'),
                    showframe = TRUE,
                    showcoastlines = TRUE)) %>%  
  plotly_build()
```

### Correlation between gold prices and World Development Indicators
To explore correlation between gold prices and global development indicators
(for example whole world's GDP), several operations on data sets were made:

1. Gold prices data set was grouped by year and stored into
`yearly_gold_prices_df` data frame
    + `gold_price_usd` column was added, which is a mean of gold USD prices
    in the morning and in the afternoon
    + Rows with `NA` values of `gold_price_usd` column were dropped
    + Whole data set was grouped by year
    + Gold price was aggregated, calculating mean gold price per year  
2. S&P Composite data set was grouped by year and stored into `yearly_sp_df`
data frame
    + Whole data set was grouped by year
    + S&P index was aggregated, calculating mean S&P index per year
3. Resulting data sets from operations 2. and 3. were merged with data set of
world's development indicators and stored into `world_joined_df` data frame
    + Subset of rows with `country_name` equal to `"World"` was selected
    from data set which only includes columns with less than 50% of `NA` values
    + Inner join by `year` column with `yearly_gold_prices_df` data set from
    operation 1. was performed. This way only observations for years with 
    known gold price were preserved
    + Left join by `year` column with `yearly_sp_df` data set from operation 2.
    was performed. This way no observations with known gold price were discarded

Bitcoin prices data set was discarded from further analysis, because 
it didn't contain data before year 2009 - Bitcoin didn't exist before
that date. Including it into resulting joined data set would result in more than
50% observations with `NA` values for Bitcoin price variable.
```{r}
yearly_gold_prices_df <- gold_prices_df %>% 
  mutate(gold_price_usd = (usd_am + usd_pm) / 2) %>%
  drop_na(gold_price_usd) %>%
  group_by(year = year(date)) %>% 
  summarize(gold_price_usd = mean(gold_price_usd))

yearly_sp_df <- sp_df %>% 
  group_by(year = year(date)) %>%
  summarize(sp_composite = mean(s_p_composite))

world_joined_df <- wdi_na_filtered_df %>%
  filter(`Country Name` == 'World') %>% 
  inner_join(yearly_gold_prices_df, by = c('Year' = 'year')) %>%
  left_join(yearly_sp_df, by = c('Year' = 'year')) %>%
  rename(`Gold price (USD)` = gold_price_usd,
         `S&P Composite` = sp_composite)
```

In the table below, all absolute correlations with gold price higher than 0.6
were presented.
```{r}
world_cor <- world_joined_df %>%
  select(3:last_col()) %>%
  cor(use = 'pairwise.complete.obs') %>%
  remove_empty()

world_gold_cor <- world_cor %>% 
  as_cordf() %>%
  focus(`Gold price (USD)`) %>% 
  arrange(desc(abs(`Gold price (USD)`))) %>% 
  filter(abs(`Gold price (USD)`) > 0.6) %>%
  rename(`Gold price (USD) correlation` = `Gold price (USD)`,
         Variable = term)

report_table(world_gold_cor, round_digits = 5)
```

## Gold price regressor
### Reducing pair-wise correlations
Since in the table of gold correlations with other variables there is more than
50 variables with absolute correlation higher than 0.6, `caret::findCorrelation`
function with `cutoff` parameter set to `0.9` was used to reduce 
pair-wise correlations. 
```{r}
vars_to_be_removed <- world_cor %>% findCorrelation()

world_gold_cor_subset <- world_joined_df %>% 
  select(3:last_col()) %>%
  select(!all_of(vars_to_be_removed)) %>%
  correlate() %>%
  focus(`Gold price (USD)`) %>%
  arrange(desc(abs(`Gold price (USD)`))) %>% 
  filter(abs(`Gold price (USD)`) > 0.6) %>%
  rename(`Gold price (USD) correlation` = `Gold price (USD)`,
         Variable = term)

correlated_variables <- world_gold_cor_subset %>% pull(Variable)
```

Most significant correlations were presented on the graph below.
```{r}
world_gold_cor_plot <- world_gold_cor_subset %>% 
  ggplot(aes(x = reorder(Variable, `Gold price (USD) correlation`), 
             y = `Gold price (USD) correlation`,
             text = paste('Variable:', Variable,
                          '<br>Gold price (USD) correlation:', 
                          round(`Gold price (USD) correlation`, digits = 5)))) + 
  geom_col() + 
  ylim(-1, 1) +
  labs(x = '') + 
  coord_flip() + 
  theme_minimal()

ggplotly(world_gold_cor_plot, tooltip = 'text')
```

### Handling missing values
Before training the model, there is a need to handle `NA` values in the data
set in some way. 

Missing values were imputed by using `missForest` library, which is based on
Random Forest algorithm. `missForest` is one of the most accurate imputation 
methods, but at the cost of slow processing time. In this case, slow processing 
time is not a problem, because input data set is relatively small.
```{r}
miss_forest <- world_joined_df %>%
  select(3:last_col()) %>%
  as.data.frame() %>%
  missForest(verbose = TRUE) 

print(miss_forest$OOBerror)
```

### Training the model
Data set was split into training and testing partitions in 75/25 ratio. 
A validation set was created using repeated cross-validation with
number of divisions equal to 2 and number of repetitions equal to 5.
```{r}
imputed_df <- miss_forest$ximp %>%
  select(all_of(correlated_variables)) %>%
  cbind(Year = world_joined_df$Year,
        `Gold price (USD)` = world_joined_df$`Gold price (USD)`)

train_index <- createDataPartition(imputed_df$`Gold price (USD)`, 
                                   p = 0.75, 
                                   list = FALSE)
train_data <- imputed_df[train_index,]
test_data <- imputed_df[-train_index,]

ctrl <- trainControl(method = "repeatedcv",
                     number = 2,
                     repeats = 5)
```

The following graph shows the similarity of the distributions 
of the training and test data. Since input data consists only of 
`r nrow(imputed_df)` observations, there is some discrepancy between training 
and testing data partitions.
```{r}
ggplot() +
  geom_density(aes(`Gold price (USD)`, fill = "Train"), train_data, alpha = 0.6) +
  geom_density(aes(`Gold price (USD)`, fill = "Test"), test_data, alpha = 0.6) +
  labs(x = "Gold price (USD)", y = "Density", fill = "Partition data set")
```

Random Forest regression model was used to predict the gold prices.
```{r}
fit <- train(`Gold price (USD)` ~ . - Year,
             data = train_data,
             method = "rf",
             trControl = ctrl)

fit
```

Two measures, `R2` and `RMSE`, were used to assess prediction accuracy.

`R2` measures the strength of the relationship between the model and 
the dependent variable on a 0 – 100% scale.

Whereas R-squared is a relative measure of fit, `RMSE` is an absolute 
measure of fit. RMSE is a good measure of how accurately the model predicts 
the response.

```{r}
prediction <- predict(fit, test_data)

post_resample <- postResample(pred = prediction,
                              obs = test_data$`Gold price (USD)`)
post_resample
```
`R2` value of `r post_resample['Rsquared']` 
and `RMSE` value of `r post_resample['RMSE']` mean that model is accurate.

The following graph shows the test set values and the model output values.
```{r}
prediction_comparison_df <- tibble(year = test_data$Year, 
                                   actual = test_data$`Gold price (USD)`,
                                   predicted = prediction)

ggplot(prediction_comparison_df, aes(x = year)) +
  geom_line(aes(y = actual, color = "Test data set")) +
  geom_line(aes(y = predicted, color = "Predicted")) +
  labs(color = "Values", x = "Year", y = "Gold price (USD)")
```

### Variable importance
```{r}
ggplot(varImp(fit))
```